---
title: "Nutrient Use in Polyploid Duckweed"
author: "Thibault Ph. Nozahic"
date: "2026-04-28"
output: pdf
---

# Nutrient omission experiment

```{r Data & Libraries import, results='hide'}

library(brms)
library(tidybayes)
library(bayesplot)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(stringr)
library(deSolve)
library(cmdstanr)
library(purrr)

#fresh weights from the nutrient omission experiment
wo <- read.table("weight_omission.txt", sep="\t", header=TRUE, dec=",") %>%
  mutate(time = round(time, 5),
         id = as.factor(id),
         time.id = as.factor(time.id)) %>%
         dplyr::select(-day,-hour,-min, -c(10:15))

#quota determined through icp-oes
qo <- read.table("quota_omission.txt", sep="\t", header=TRUE, dec=",") %>%
  mutate(id = as.factor(id),
         P = round(P,3))

ic <- read.table("ic_omission.txt", sep="\t", header=TRUE, dec=",") %>%
  mutate(id = as.factor(id))

#determining q0 & qend for each nutrient:ploidy couple
Q0 <- bind_rows(
  qo %>%
    filter(time.id == 0, P != 0) %>%
    group_by(medium, ploidy) %>%
    summarise(Q0 = mean(P), .groups = "drop") %>%
    mutate(medium = "P"),

  qo %>%
    filter(time.id == 0, N != 0) %>%
    group_by(medium, ploidy) %>%
    summarise(Q0 = mean(N), .groups = "drop") %>%
    mutate(medium = "N")
)

Qend <- qo %>%
  filter(time.id == 9) %>%
  mutate(
    Qend = case_when(
      medium == "P" ~ P,
      medium == "N" ~ N
    )
  ) %>%
  group_by(medium, ploidy) %>%
  summarise(Qend = mean(Qend), .groups = "drop")

wo <- wo %>%
  left_join(Q0, by = c("ploidy","medium")) 

```

`wo` for weight omission, `qo` for quota omission & `ic` for ion chromatography

The Droop model (Tilman, 1977) is simplified to yield the two simplified equations:

$$\frac{dN}{dt} = N \times (r \times (1 - \frac{q_{min}}Q) - m) \\ \frac{dQ}{dt} = - r \times (Q - q_{min})$$

thus: $$Q(t) = q_{min} + (Q_0 - q_{min}) \times e^{-rt} \\ N(t) = \frac{e^{-mt} \times N_0 \times Q_0}{q_{min} + (Q_0 - q_{min}) \times e^{-rt}}$$

$$Growth = \frac{N(t)}{N_0} \\ ln(Growth) = -m \times t + ln(Q_0) - ln(q_{min} + (Q_0 - q_{min}) \times e^{-rt})$$

to formulate this equation using the probabilistic formulation of *brms*

$$lnGrowth\ \mathtt{\sim}\ Normal(\mu_N, \sigma) \\ lnGrowth =  -mortality \times time + log(Q_0) - log(q_{min} + (Q_0 - q_{min}) \times  e^{-r\ time})\\ mortality + qmin + r\ \mathtt{\sim}\ \ 0 + ploidy:medium$$ the latter means that the three parameters (`mortality`, `qmin` & `r`) to fit are linked to the specified `ploidy` and `medium`

```{r Building the model, priors and fitting, results='hide'}

#building the model, setting the priors & fitting

rubbish <- bf(lngrowth ~ 
                 -mortality * time + log(Q0) 
              - log(qmin + (Q0 - qmin) * exp(-r * time)),
              
              r + mortality + qmin ~ 0,
              
              nl = TRUE)

#stick with this but we can try with (r ~ 0 + ploidy) only
lnGrowth <- bf(lngrowth ~
                  -mortality * time + log(Q0) 
               - log(qmin + (Q0 - qmin) * exp(-r * time)),
               
               qmin + mortality + r ~ 0 + ploidy:medium,
               
               nl = TRUE)

#with r ~ 0 + ploidy only
lnGrowthrploidy <- bf(lngrowth ~
                  -mortality * time + log(Q0) 
               - log(qmin + (Q0 - qmin) * exp(-r * time)),
               
               qmin + mortality ~ 0 + ploidy:medium,
               r ~ 0 + ploidy,
               
               nl = TRUE)

#probably bad but still question
lnGrowth1 <- bf(lngrowth ~
                  -mortality * time + log(Q0) 
               - log(qmin + (Q0 - qmin) * exp(-r * time)),
               
               qmin ~ 0 + ploidy:medium,
               r + mortality ~ 0 + ploidy,
               
               nl = TRUE)

priors <- prior(normal(0.5, 0.1), nlpar = "qmin", lb = 0) +
          prior(normal(0.01, 0.05), nlpar = "r", lb = 0) +
          prior(normal(0.02, 0.05), nlpar = "mortality", lb = 0) +
          prior(cauchy(0, 0.5), class = "sigma")

priors <- prior(normal(2, 5), nlpar = "qmin", lb = 0) +
          prior(normal(0, 0.1), nlpar = "r", lb = 0) +
          prior(normal(0, 0.1), nlpar = "mortality", lb = 0) +
          prior(cauchy(0, 1), class = "sigma")

options(mc.cores = parallel::detectCores())

fitting <- brm(lnGrowth,
               data = wo,
               prior= priors,
               iter = 5000, chains = 8,
               backend="cmdstanr")

#quick check of the results
saveRDS(fitting, "fittingwithpriors0.5_0.5-0_0.1-0_0.1-0_1.modelqminmortalityandr.RDS")
fitting <- readRDS("C:/Users/tozah/Desktop/Internship Ghent University/Protocols & Experiments/fittingwithpriors0.5_0.5-0_0.1-0_0.1-0_1.modelqminmortalityandr.RDS")

summary(fitting)

summfitting <- summary(posterior)
plot(fitting)

writexl::write_xlsx(summfitting, "fitting.xlsx")

posterior <- as_draws_df(fitting) %>%
  mutate(
    `Δqmin NO3` = (`b_qmin_ploidydip:mediumN` - `b_qmin_ploidytet:mediumN`),
    `Δqmin PO4` = (`b_qmin_ploidydip:mediumP` - `b_qmin_ploidytet:mediumP`),
    `Δr w/o NO3` = (`b_r_ploidydip:mediumN` - `b_r_ploidytet:mediumN`),
    `Δr w/o PO4` = (`b_r_ploidydip:mediumP` - `b_r_ploidytet:mediumP`),
    `Δmortality w/o NO3` = (`b_mortality_ploidydip:mediumN` - `b_mortality_ploidytet:mediumN`),
    `Δmortality w/o PO4` = (`b_mortality_ploidydip:mediumP` - `b_mortality_ploidytet:mediumP`))


#first version for r + mortality + qmin ~ 0 + ploidy:medium
#post is only used to plot Q(t)
post <- as_draws_df(fitting) %>%
  select(.draw, matches("^b_r_"), matches("^b_qmin_")) %>%
  pivot_longer(
    cols = - .draw,
    names_to = "term",
    values_to = "value"
  ) %>%
  mutate(
    ploidy = ifelse(grepl("ploidydip", term), "dip", "tet"),
    medium = ifelse(grepl("mediumN", term), "N", "P"),
    param  = case_when(
      grepl("^b_r_", term) ~ "r",
      grepl("^b_qmin_", term) ~ "qmin"
    )
  ) %>%
  select(.draw, ploidy, medium, param, value) %>%
  pivot_wider(names_from = param, values_from = value)

#second version for r + mortality ~ 0 + ploidy, qmin ~ 0 + ploidy:medium
post_r <- as_draws_df(fitting) %>%
  select(.draw, matches("^b_r_")) %>%
  pivot_longer(cols = - .draw) %>%
  mutate(
    ploidy = ifelse(grepl("ploidydip", name), "dip", "tet")
  ) %>%
  rename(r = value)

post_qmin <- as_draws_df(fitting) %>%
  select(.draw, matches("^b_qmin_")) %>%
  pivot_longer(cols = - .draw) %>%
  mutate(
    ploidy = ifelse(grepl("ploidydip", name), "dip", "tet"),
    medium = ifelse(grepl("mediumN", name), "N", "P")
  ) %>%
  rename(qmin = value)

```

```{r Plot of the lnGrowth(t)}

label_param <- function(var) {
  case_when(
    grepl("qmin", var)      ~ "Δ q[min]",
    grepl("mortality", var) ~ "Δ mortality",
    grepl("r", var)         ~ "Δ r",
    grepl("V", var)         ~ "Δ V",
    grepl("K", var)         ~ "Δ K"
  )
}

label_medium <- function(var) {
  case_when(
    grepl("NO3|N", var) ~ "NO\u2083",
    grepl("PO4|P", var) ~ "PO\u2084"
  )
}

label_medium_wo <- function(var) {
  case_when(
    grepl("N", var) ~ "w/o NO\u2083",
    grepl("P", var) ~ "w/o PO\u2084"
  )
}


#new data set in which we'll put the estimated points
est <- wo %>%
  distinct(ploidy, medium) %>%
  left_join(Q0, by = c("ploidy", "medium")) %>%
  tidyr::crossing(time = seq(min(wo$time), max(wo$time), length.out = 200))

#fitted points using the previous fitting
pred <- fitted(fitting,
               newdata = est,
               re_formula = NA,
               scale = "response") %>%
  as.data.frame()

est <- wo %>%
  distinct(ploidy, medium) %>%
  left_join(Q0, by = c("ploidy", "medium")) %>%
  tidyr::crossing(time = seq(min(wo$time), max(wo$time), length.out = 200))

est <- bind_cols(est, pred) %>%
  mutate(
    medium = c("NO3 omission", "PO4 omission")[match(medium, c("N", "P"))])

#plotting with Q2.5, Q97.5
ggplot(wo %>% mutate(medium = c("NO3 omission", "PO4 omission")[match(medium, c("N", "P"))]), aes(time, lngrowth)) +
  geom_point(aes(color = ploidy), size = 1.5, alpha = 0.7) +
  geom_line(data = est,
            aes(y = Estimate, color = ploidy),
            linewidth = 1) +
  geom_ribbon(data = est,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5, fill = ploidy),
              alpha = 0.2, color = NA) +
  scale_color_manual(values = c("dip" = "royalblue", "tet" = "deeppink3")) +
  scale_fill_manual(values  = c("dip" = "dodgerblue", "tet" = "deeppink")) +
  ylab("lnGrowth") +
  xlab("Time (days)") +
  facet_wrap(~ medium) + 
  theme_bw()

#saving the plot
ggsave(filename="growthfinal1.png")

```

```{r Plot of Q(t)}

#first version for r + mortality + qmin ~ 0 + ploidy:medium
est_Q <- est %>%
  left_join(post, by = c("ploidy", "medium")) %>%
  mutate(Q = qmin + (Q0 - qmin) * exp(-r * time))

#second version for r + mortality ~ 0 + ploidy, qmin ~ 0 + ploidy:medium
post_qmin_r <- post_qmin %>%
  left_join(post_r, by = c(".draw", "ploidy"))

est_Q <- est %>%
  crossing(.draw = unique(post_qmin_r$.draw)) %>%
  left_join(post_qmin_r, by = c("ploidy", "medium", ".draw")) %>%
  mutate(Q = qmin + (Q0 - qmin) * exp(-r * time))

#same for both we extract the median, Q2.5, Q97.5
est_summary <- est_Q %>%
  group_by(ploidy, medium, time) %>%
  summarise(
    Estimate = median(Q),
    Q2.5     = quantile(Q, 0.025),
    Q97.5    = quantile(Q, 0.975),
    .groups = "drop"
  )

#the points q0 and qend together by medium and ploidy with correct time labels
q_points <- Q0 %>%
  left_join(Qend, by = c("medium", "ploidy")) %>%
  tidyr::pivot_longer(
    cols = c(Q0, Qend),
    names_to = "label",
    values_to = "Q"
  ) %>%
  mutate(time = ifelse(label == "Q0", 0, 21))

#plot
ggplot() +
  geom_point(data = q_points,
             aes(time, Q, color = ploidy, shape = label),
             size = 3) +
  geom_line(data = est_summary,
            aes(time, Estimate, color = ploidy),
            linewidth = 1) +
  geom_ribbon(data = est_summary,
              aes(time, ymin = Q2.5, ymax = Q97.5, fill = ploidy),
              alpha = 0.2, color = NA) +
  scale_color_manual(values = c("dip" = "royalblue", "tet" = "deeppink3")) +
  scale_fill_manual(values = c(dip = "dodgerblue", tet = "deeppink")) +
  facet_wrap(~ medium, scales = "free_y") +
  theme_bw()

#saving the plot
ggsave(filename="quotastoragefinalfree_y.png")

```

```{r Plot the posterior parameters distribution}

#first version with r and mortality ~ 0 + ploidy:medium
df_long1 <- posterior %>%
  dplyr::select(-.chain, -.iteration, -.draw, -sigma, -lprior, -lp__, -`Δqmin NO3`, -`Δqmin PO4`, -`Δr w/o NO3`, -`Δr w/o PO4`, -`Δmortality w/o NO3`, -`Δmortality w/o PO4`) %>%
  tidyr::pivot_longer(everything(),
                      names_to = "variable",
                      values_to = "value") %>%
  dplyr::mutate(
    param  = stringr::str_extract(variable, "qmin|mortality|(?<=_)r(?=_)"),
    ploidy = stringr::str_extract(variable, "dip|tet"),
    medium = stringr::str_extract(variable, "(?<=medium)[NP]")
  ) %>%
  # drop any leftover posterior columns that aren't qmin/r/mortality x ploidy x
  # medium terms -- this is what produced the spurious "NA"/"NA NA" panel
  tidyr::drop_na(param, ploidy, medium) %>%
  dplyr::mutate(
    # qmin is estimated in both media; r and mortality can only be estimated
    # from the medium where the *other* nutrient was omitted
    medium_label = dplyr::case_when(
      param == "qmin" & medium == "N" ~ "NO3",
      param == "qmin" & medium == "P" ~ "PO4",
      medium == "N" ~ "NO3 omitted",
      medium == "P" ~ "PO4 omitted"
    ),
    label = paste0(ploidy, " ", medium_label),
    # tet/dip pairs adjacent within each medium, for easier ploidy comparison
    label = factor(label, levels = c("dip PO4", "tet PO4", "dip NO3", "tet NO3",
                                      "dip PO4 omitted", "tet PO4 omitted",
                                      "dip NO3 omitted", "tet NO3 omitted"))
  )


df_long1 %>%
  ggplot(aes(y = label, x = value, fill = label)) +
  geom_violin(quantiles = c(0.1, 0.5, 0.9), 
              quantile.linetype = 1, quantile.colour = "grey95") +
  
  scale_fill_manual(values = c(`dip NO3` = "dodgerblue",
                               `tet NO3` = "deeppink",
                               `dip PO4` = "royalblue",
                               `tet PO4` = "deeppink3",
                               `dip NO3 omitted` = "dodgerblue",
                               `tet NO3 omitted` = "deeppink",
                               `dip PO4 omitted` = "royalblue",
                               `tet PO4 omitted` = "deeppink3")) +
  facet_wrap(~ param, scales = "free", ncol = 1,
             labeller = as_labeller(c(qmin = "\u0394qmin (g/kg)",
                                       r = "\u0394r (day\u207b\u00b9)",
                                       mortality = "\u0394mortality (day\u207b\u00b9)"))) +
  labs(x = "Posterior estimate", y = NULL, fill = NULL) +
  theme_bw() +
  theme(legend.position = "none")

ggsave("parameeeeetersfinal.png")


differences <- posterior %>%
  dplyr::select(`Δqmin NO3`, `Δqmin PO4`, `Δr w/o NO3`, `Δr w/o PO4`, `Δmortality w/o NO3`, `Δmortality w/o PO4`) %>%
  tidyr::pivot_longer(everything(),
                      names_to = "variable",
                      values_to = "value") %>%
  mutate(medium = ifelse(grepl("N", variable), "N", "P"),
         param = case_when(grepl("qmin", variable) ~ "qmin",
                           grepl("mortality", variable) ~ "mortality",
                           grepl("r", variable) ~ "r"))

differences %>%
  ggplot(aes(y = variable, x = value)) +
  geom_violin(fill = "grey50", color = "grey30", quantiles = c(0.1, 0.5, 0.9), 
              quantile.linetype = 1, quantile.colour = "black") +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.7) +
  labs(x = "Posterior estimate", y = NULL, fill = NULL) +
  facet_wrap(~ param, nrow = 3, scales = "free") +
  theme_bw()

#other representation but with r and mortality independent from nutrient so only ploidy-related
df_long <- posterior %>%
  dplyr::select(-.chain, -.iteration, -.draw, -sigma, -lprior, -lp__) %>%
  tidyr::pivot_longer(everything(),
                      names_to = "variable",
                      values_to = "value") %>%
  dplyr::mutate(
    param  = stringr::str_extract(variable, "qmin|mortality|(?<=_)r(?=_)"),
    ploidy = stringr::str_extract(variable, "dip|tet"),
    medium = stringr::str_extract(variable, "(?<=medium)[NP]"),
    label  = paste0(ploidy, ".", ifelse(is.na(medium), "", medium))
  )

df_quantiles <- df_long %>%
  summarise(
    q5 = quantile(value, 0.05),
    q95 = quantile(value, 0.95),
    .by = c(label, param)
  )

df_long %>%
  ggplot(aes(x = label, y = value, fill = label)) +
  geom_violin() +

  geom_segment(
    data = df_quantiles,
    aes(
      x = label,
      xend = label,
      y = q5,
      yend = q95,
    ),
    color = "white",
    linewidth = 0.7,
    inherit.aes = FALSE
  ) +
  
  stat_summary(fun = median, geom = "point", color = "white", size = 1) +
  scale_fill_manual(values = c(dip.N = "dodgerblue", 
                               dip.P = "royalblue", 
                               tet.N = "deeppink", 
                               tet.P = "deeppink3",
                               dip. = "dodgerblue",
                               tet. = "deeppink")) +
  facet_wrap(~ param, scales = "free", ncol = 1) +
  facet_grid(~ param, scales = "free", space = "free") +
  labs(x = "Posterior estimate", y = NULL, fill = NULL) +
  theme_bw() +
  theme(legend.position = "hidden")

#saving the plot
ggsave(filename="diffparamfinalfinal.png")

```

# Nutrient uptake experiment

```{r Data counterfeit & import, results='hide'}
library(brms)
library(tidybayes)
library(bayesplot)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(stringr)
library(deSolve)
library(cmdstanr)
library(purrr)

##==============================================================================

##==============================================================================
S_function <- function(time, state, parms) {
  S  <- state[["S"]]
  N0 <- parms[["N0"]]
  r <- parms[["r"]]
  V  <- parms[["V"]]
  K  <- parms[["K"]]

  dS <- -N0 * exp(r * time) * V * S / (K + S)
  list(c(dS))
}

conditions <- tibble(
condition = c("dip0.25P", "dip0.1P", "dip0.02P", "dip0.2N", "dip0.1N", "dip0.02N", "dip0.005N","tet0.25P", "tet0.1P", "tet0.02P", "tet0.2N", "tet0.1N", "tet0.02N", "tet0.005N"
),
ploidy = c(
"dip","dip","dip","dip","dip","dip","dip",
"tet","tet","tet","tet","tet","tet","tet"),
nutrient = c(
"P","P","P","N","N","N","N",
"P","P","P","N","N","N","N"),
V = c(
1,1,1,5,5,5,5,
2,2,2,6,6,6,6),
K = c(
3,2,4,60,76,63,70,
2,3,5,58,61,61,67),
r = c(
0.2,0.2,0.2,0.2,0.2,0.2,0.2,
0.1,0.1,0.1,0.1,0.1,0.1,0.1),
S_N = c(
550, 550, 550, 110, 55, 11, 2.5,
550, 550, 550, 110, 55, 11, 2.5),
S_P = c(
12.5, 5, 1, 50, 50, 50, 50,
12.5, 5, 1, 50, 50, 50, 50),
N0 = 1
)

time <- seq(from = 0, to = 10, by = 10)

counterfeit <- function(V, K, r, N0, S_N, S_P, nutrient) {
  S0 <- if (nutrient == "N") S_N else S_P

  parms <- c(N0 = N0, r = r, V = V, K = K)
  state <- c(S = S0)

  deSolve::ode(y = state, times = time, func = S_function, parms = parms) %>%
    as.data.frame()
}

output <- conditions %>%
  mutate(output = pmap(
    list(V, K, r, N0, S_N, S_P, nutrient),
    counterfeit
  ))

noise <- 0

out <- output %>%
  unnest(output) %>%
  mutate(
    R_true   = S,
    R_obs    = R_true + rnorm(n(), mean = 0, sd = noise),
    logR_obs = log(R_obs),
    R0       = if_else(nutrient == "N", S_N, S_P)
  ) %>%
  select(condition, ploidy, nutrient, time, R_true, R_obs, logR_obs,
         R0, N0, r, V, K) %>%
  filter(time > 0)

##==============================================================================

#import the real data

nu <- read.table("nutrient_uptake_incomplete.txt", header = T, sep = "\t", dec = ",") %>%
  select(id, ploidy, medium, rep, observer, 
         time, weight0, weight_pot_lid, weight_pot_lid_med, total_weight_begin, total_weight_end, fw.end, growth, lngrowth,
         mg.l.SO4, mg.l.NO3, mg.l.PO4) %>%
  mutate(condition = interaction(ploidy, medium, sep = ";"),
         volume = (weight_pot_lid_med - weight_pot_lid)/1000,
         conc.weight0 = weight0 * volume,
         conc.fw = fw.end * volume,
         growth = conc.fw / conc.weight0,
         lngrowth = log(growth),
         r = ifelse(ploidy == "dip", 0.06622310, 0.05511983),
         N0 = conc.weight0,
         time = round(time, 5)) %>%
  rename(
    NO3 = mg.l.NO3,
    PO4 = mg.l.PO4
  ) %>%
  pivot_longer(
    cols = c("NO3", "PO4"),
    values_to = "S_unscaled",
    names_to = "nutrient"
  ) %>%
  filter(
    (grepl("N", medium) & nutrient == "NO3") |
    (grepl("P", medium) & nutrient == "PO4") |
    (medium == "0.2N")) %>%
  mutate(censored = ifelse(is.na(S_unscaled), -1, 0),
         S_unscaled = ifelse(is.na(S_unscaled), 0.2, S_unscaled),
         S_scaled = S_unscaled * (96.06/mg.l.SO4),
         logS = log(S_scaled)) %>%
  group_by(medium, nutrient) %>%
  mutate(S0 = mean(S_scaled[time == 0], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(
    mednut = paste0(nutrient, "::", medium),
    ploinut = paste0(ploidy, "::", medium, "::", nutrient))

S0 <- nu %>%
  filter(time == 0) %>%
  select(medium, nutrient, S0) %>%
  distinct() %>%
  as.data.frame()

ggplot(nu, aes(medium, S_scaled, color = ploidy)) +
  geom_point(size = 1.5, alpha = 0.7) +
  scale_color_manual(values = c("dip" = "royalblue", "tet" = "deeppink3")) +
  scale_fill_manual(values  = c("dip" = "dodgerblue", "tet" = "deeppink")) +
  facet_wrap(~ nutrient, scales = "free") +
  theme_bw()

ggplot(nu%>%filter(medium == "0.005N", nutrient == "NO3"), aes(ploidy, logS, color = ploidy)) +
  geom_point(size = 1.5, alpha = 0.7) +
  scale_color_manual(values = c("dip" = "royalblue", "tet" = "deeppink3")) +
  scale_fill_manual(values  = c("dip" = "dodgerblue", "tet" = "deeppink")) +
  theme_bw()


```

```{r C++ Stan code, results='hide'}

ode <- "
vector dXdt(real time,
            vector X,
            array[] real theta) {

  real V = theta[1];
  real logK = theta[2];
  real N0   = theta[3];
  real r    = theta[4];
  
  real K = exp(logK);

  vector[1] dXdt;
  dXdt[1] = -N0 * exp(r * time) * V / (K + exp(X[1]));
  return dXdt;
}

real Rintegrator(real time, real S0, real V, real logK, real N0, real r) {
  vector[1] y0;
  array[4] real alpha;

  y0[1] = S0;

  alpha[1] = V;
  alpha[2] = logK;
  alpha[3] = N0;
  alpha[4] = r;

  real t0 = 0.0;
  real ts = time;
  if (ts <= t0) {
    ts = t0 + 1e-9;
  }

  array[1] vector[1] y;
  y = ode_rk45(dXdt, y0, t0, rep_array(ts, 1), alpha);

  return y[1][1];
}
"

odelog <- "
vector dXdt(real time,
            vector X,
            array[] real theta) {

  real logV = theta[1];
  real logK = theta[2];
  real N0   = theta[3];
  real r    = theta[4];

  real V = exp(logV);
  real K = exp(logK);

  vector[1] dXdt;
  dXdt[1] = -N0 * exp(r * time) * V / (K + exp(X[1]));
  return dXdt;
}

real Rintegratorlog(real time, real S0, real logV, real logK, real N0, real r) {
  vector[1] y0;
  array[4] real alpha;

  y0[1] = log(S0);

  alpha[1] = logV;
  alpha[2] = logK;
  alpha[3] = N0;
  alpha[4] = r;

  real t0 = 0.0;
  real ts = time;
  if (ts <= t0) {
    ts = t0 + 1e-9;
  }

  array[1] vector[1] y;
  y = ode_rk45(dXdt, y0, t0, rep_array(ts, 1), alpha);

  return exp(y[1][1]);
}
"

odelogS <- "
  vector dSdt(real t,
              vector S,
              array[] real theta) {

    real V = theta[1];
    real K   = theta[2];
    real N0    = theta[3];   
    real r    = theta[4];   
    vector[1] dSdt;
    dSdt[1] = -1*N0*exp(r*t) * V * S[1] / (K + S[1]);
    return dSdt;
    }

real Rintegrator(real t, real S0, real V, real K, real N0, real r) {
  vector[1] y0;
  array[4] real alpha;// Parameters
  
  y0[1] = S0;

  alpha[1] = V; 
  alpha[2] = K;
  alpha[3] = N0;
  alpha[4] = r;

  array[1] vector[1] y;
  y = ode_rk45(dSdt,
                y0, 0, rep_array(t, 1),
                //0.01, 0.01, 100,
                alpha); 
  return y[1][1];
}
"
```

```{r BRMS run, results='hide'}

options(mc.cores = parallel::detectCores())

lilgrowth <- bf(growth ~ exp(r * time/24),  r ~ 0 + ploidy,  nl = TRUE)
r <- prior(normal(0,1), nlpar = r, lb = 0)
fitlilgrowth <- brm(lilgrowth, data = nu, prior = r, backend = "cmdstanr")
summary(fitlilgrowth)
fitidoodie <- as_draws_df(fitlilgrowth)
summary(fitidoodie)

bf_logS <- bf(
  logS | cens(censored) ~
    log(Rintegrator(time/24, S0, V, K, N0, r))+o,
  nlf(V ~ exp(logV)),
  nlf(K ~ exp(logK)),
  logV + logK ~ 0 + ploidy:nutrient,
  o ~ 0 + (1|observer),
  nl = TRUE
)

get_prior(bf_logS, data = nu)
priors <- c(
  prior(normal(6, 2), class = "b", nlpar = "logV"),
  prior(normal(2, 2), class = "b", nlpar = "logK"),
  prior(cauchy(0, 0.1), class = "sd", nlpar = "o"),
  prior(cauchy(0, 1), class = "sigma")
)

fodebest <- brm(
  formula = bf_logS,
  data = nu,
  prior = priors,
  stanvars = stanvar(scode = odelogS, block = "functions"),
  backend = "cmdstanr"
)

summary(fodebest)
fodebest$fit
plot(fodebest)

saveRDS(fodebest, file = "fodebestlast.RDS")
fode <- readRDS("C:/Users/tozah/Desktop/Internship Ghent University/Protocols & Experiments/fodedata6.1august.RDS")

expose_functions(fodebest, vectorize = T)

posteriorVK <- as_draws_df(fodebest)

est1 <- nu %>%
  filter(!is.na(N0)) %>%
  distinct(ploidy, medium, nutrient, N0, r, S0, id) %>%
  tidyr::crossing(time = seq(1e-9, 25, length.out = 10)) %>%
  mutate(
    ploidy   = factor(ploidy,   levels = c("dip", "tet")),
    nutrient = factor(nutrient, levels = c("NO3", "PO4")),
    condition = interaction(ploidy, medium, sep = "::"),
    id = factor(id)
  )

#fitted points using the previous fitting

fodepred <- fitted(
  fodebest,
  newdata = est1,
  re_formula = NA
)

fodepredback <- fodepred %>% 
  as.data.frame() %>%
  mutate(
    est1, 
    across(c(1, 3, 4), exp),
    mednut = paste0(nutrient, "::", medium),
    ploinut = paste0(ploidy, "::", medium, "::", nutrient)
  )


ggplot(nu, aes(time, S_scaled, group = id)) +
  geom_point(aes(color = ploidy), size = 1.5, alpha = 0.7) +
  geom_line(data = fodepredback,
            aes(y = Estimate, color = ploidy),
            linewidth = 0.5) +
  geom_ribbon(data = fodepredback,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5, 
                  color = ploidy, fill = ploidy),
              alpha = 0.1, color = NA) +
  scale_color_manual(values = c("dip" = "royalblue", "tet" = "deeppink3")) +
  scale_fill_manual(values  = c("dip" = "dodgerblue", "tet" = "deeppink")) +
  facet_wrap(~ mednut, scales = "free_y", ncol = 4) +
  theme_bw()

ggsave("besttry88889log.png")

```

```{r }


ggplot(nu, aes(time, S_scaled)) +
  geom_point(aes(color = ploidy), size = 1.5, alpha = 0.7) +
  scale_color_manual(values = c("dip" = "royalblue", "tet" = "deeppink3")) +
  scale_fill_manual(values  = c("dip" = "dodgerblue", "tet" = "deeppink")) +
  facet_wrap(~ nutrient, scales = "free_y") +
  theme_bw()

geom_line(data = estS_summary,
            aes(y = Estimate, color = ploidy),
            linewidth = 1) +
  geom_ribbon(data = estS_summary,
              aes(y = Estimate, ymin = Q2.5, ymax = Q97.5, fill = ploidy),
              alpha = 0.2, color = NA)

posteriorVK__ <- posteriorVK %>%
  mutate(
    across(c(1:8), exp),
    dVN = `b_logV_ploidydip:nutrientNO3` - `b_logV_ploidytet:nutrientNO3`,
    dVP = `b_logV_ploidydip:nutrientPO4` - `b_logV_ploidytet:nutrientPO4`,
    dKN = `b_logK_ploidydip:nutrientNO3` - `b_logK_ploidytet:nutrientNO3`,
    dKP = `b_logK_ploidydip:nutrientPO4` - `b_logK_ploidytet:nutrientPO4`
  )

names(posteriorVK)

df_longVK <- posteriorVK__ %>%
  dplyr::select(-.chain, -.iteration, -.draw, -sigma, -lprior, -lp__, -dVN, -dVP, -dKN, -dKP, -sd_observer__o_Intercept, -`r_observer__o[Frederik,Intercept]`, -`r_observer__o[Thibault,Intercept]`) %>%
  tidyr::pivot_longer(everything(),
                      names_to = "variable",
                      values_to = "value") %>%
  dplyr::mutate(
    param  = stringr::str_extract(variable, "V|K|(?<=_)r(?=_)"),
    ploidy = stringr::str_extract(variable, "dip|tet"),
    medium = stringr::str_extract(variable, "NO3|PO4"),
    label  = paste0(ploidy, ".", medium),
    parlab = paste0(param, ".", medium)
  )

df_longVK %>%
  ggplot(aes(y = label, x = value, fill = label)) +
  geom_violin(quantiles = c(0.1, 0.5, 0.9), 
              quantile.linetype = 1, quantile.colour = "grey95") +
  
  scale_fill_manual(values = c(dip.NO3 = "dodgerblue", 
                               dip.PO4 = "royalblue", 
                               tet.NO3 = "deeppink", 
                               tet.PO4 = "deeppink3")) +
  facet_wrap(~ parlab, scales = "free", ncol = 1) +
  labs(x = "Posterior estimate", y = NULL, fill = NULL) +
  theme_bw() +
  theme(legend.position = "right")

ggsave("trybest888.png")

differences <- posteriorVK__ %>%
  dplyr::select(dVN, dVP, dKN, dKP) %>%
  tidyr::pivot_longer(everything(),
                      names_to = "variable",
                      values_to = "value") %>%
  mutate(medium = ifelse(grepl("NO3", variable), "NO3", "PO4"),
         param = case_when(grepl("V", variable) ~ "V",
                           grepl("K", variable) ~ "K"))


names(posteriorVK__)


differences %>%
  ggplot(aes(y = variable, x = value)) +
  geom_violin(fill = "grey50", color = "grey30", quantiles = c(0.1, 0.5, 0.9), 
              quantile.linetype = 1, quantile.colour = "black") +
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.7) +
  labs(x = "Posterior estimate", y = NULL, fill = NULL) +
  facet_wrap(~ param, nrow = 3, scales = "free") +
  theme_bw()

ggsave("besttry888.png")

```

```{r Finding R* and consumption vectors}

#Finding R*

Rpost <- as_draws_df(fitting)

RpostVK <- as_draws_df(fodebest)

R_etoile_post <- data.frame()
R_etoile_post <- Rpost %>%
  left_join(RpostVK, by = join_by(.draw)) %>%
  mutate(
    K_dip_NO3 = exp(`b_logK_ploidydip:nutrientNO3`),
    K_tet_NO3 = exp(`b_logK_ploidytet:nutrientNO3`),
    K_dip_PO4 = exp(`b_logK_ploidydip:nutrientPO4`),
    K_tet_PO4 = exp(`b_logK_ploidytet:nutrientPO4`),

    V_dip_NO3 = exp(`b_logV_ploidydip:nutrientNO3`),
    V_tet_NO3 = exp(`b_logV_ploidytet:nutrientNO3`),
    V_dip_PO4 = exp(`b_logV_ploidydip:nutrientPO4`),
    V_tet_PO4 = exp(`b_logV_ploidytet:nutrientPO4`),

    R_min_dip_NO3 = (`b_r_ploidydip:mediumN` * `b_qmin_ploidydip:mediumN` * `K_dip_NO3` * `b_mortality_ploidydip:mediumN`) /
                    (`V_dip_NO3` * (`b_r_ploidydip:mediumN` - `b_mortality_ploidydip:mediumN`) -
                     `b_r_ploidydip:mediumN` * `b_qmin_ploidydip:mediumN` * `b_mortality_ploidydip:mediumN`),
    R_min_tet_NO3 = (`b_r_ploidytet:mediumN` * `b_qmin_ploidytet:mediumN` * `K_tet_NO3` * `b_mortality_ploidytet:mediumN`) /
                    (`V_tet_NO3` * (`b_r_ploidytet:mediumN` - `b_mortality_ploidytet:mediumN`) -
                     `b_r_ploidytet:mediumN` * `b_qmin_ploidytet:mediumN` * `b_mortality_ploidytet:mediumN`),
    R_min_dip_PO4 = (`b_r_ploidydip:mediumP` * `b_qmin_ploidydip:mediumP` * `K_dip_PO4` * `b_mortality_ploidydip:mediumP`) /
                    (`V_dip_PO4` * (`b_r_ploidydip:mediumP` - `b_mortality_ploidydip:mediumP`) -
                     `b_r_ploidydip:mediumP` * `b_qmin_ploidydip:mediumP` * `b_mortality_ploidydip:mediumP`),
    R_min_tet_PO4 = (`b_r_ploidytet:mediumP` * `b_qmin_ploidytet:mediumP` * `K_tet_PO4` * `b_mortality_ploidytet:mediumP`) /
                    (`V_tet_PO4` * (`b_r_ploidytet:mediumP` - `b_mortality_ploidytet:mediumP`) -
                     `b_r_ploidytet:mediumP` * `b_qmin_ploidytet:mediumP` * `b_mortality_ploidytet:mediumP`))

R_tidy <- R_etoile_post %>%
  select(.draw,
         R_min_dip_NO3,
         R_min_tet_NO3,
         R_min_dip_PO4,
         R_min_tet_PO4) %>%
  pivot_longer(
    cols = starts_with("R_min"),
    names_to = c("ploidy", "nutrient"),
    names_pattern = "R_min_(.*)_(.*)",
    values_to = "R_min"
  )

R_summary <- R_tidy %>%
  group_by(ploidy, nutrient) %>%
  summarise(
    median = median(R_min, 0.50, na.rm = TRUE),
    p10 = quantile(R_min, 0.10, na.rm = TRUE),
    p90 = quantile(R_min, 0.90, na.rm = TRUE),
    .groups = "drop"
  )

R_points <- R_etoile_post %>%
  select(.draw,
         R_min_dip_NO3, R_min_dip_PO4,
         R_min_tet_NO3, R_min_tet_PO4) %>%
  pivot_longer(
    cols = starts_with("R_min"),
    names_to = c("ploidy", "nutrient"),
    names_pattern = "R_min_(.*)_(.*)",
    values_to = "R_min") %>%
  pivot_wider(
    names_from = nutrient,
    values_from = R_min,
    names_prefix = "R_min_") %>%
  filter(R_min_PO4 <= 0.00015,
         R_min_NO3 <= 0.003)

summary(posteriorVK__)

posteriorVK_summary <- summary(posteriorVK__) %>%
  as_tibble() %>%
  select(variable, median)

VK_medians <- posteriorVK_summary %>%
  filter(str_detect(variable, "b_log")) %>%
  mutate(
    param    = str_extract(variable, "(?<=b_log)(V|K)"),
    ploidy   = str_extract(variable, "(?<=ploidy)(dip|tet)"),
    nutrient = str_extract(variable, "(?<=nutrient)(NO3|PO4)"),
    median   = median   # undo log
  ) %>%
  select(param, ploidy, nutrient, median) %>%
  pivot_wider(names_from = param, values_from = median)

nu_medians <- nu %>%
  group_by(ploidy, nutrient) %>%
  summarise(
    r  = median(r),
    N0 = median(N0),
    S0 = max(S0),
    .groups = "drop"
  )

params <- VK_medians %>%
  left_join(nu_medians, by = c("ploidy", "nutrient"))
t <- 0

consumption_vectors <- params %>%
  mutate(
    dSdt = -N0 * exp(r * t) * (V * S0) / (S0 + K)
  ) %>%
  # reshape to NO3/PO4 pairs
  select(ploidy, nutrient, dSdt) %>%
  pivot_wider(names_from = nutrient, values_from = dSdt) %>%
  mutate(
    norm  = sqrt(NO3^2 + PO4^2),
    u_NO3 = NO3 / norm,
    u_PO4 = PO4 / norm,
    slope = NO3 / PO4,
    angle = atan2(PO4, NO3)
  )

arrow_df <- consumption_vectors %>%
  mutate(
    x_start = ifelse(ploidy == "dip", R_summary %>% filter(ploidy=="dip", nutrient=="NO3") %>% pull(median), R_summary %>% filter(ploidy=="tet", nutrient=="NO3") %>% pull(median)),
    y_start = ifelse(ploidy == "dip", R_summary %>% filter(ploidy=="dip", nutrient=="PO4") %>% pull(median), R_summary %>% filter(ploidy=="tet", nutrient=="PO4") %>% pull(median)),
    x_end = x_start + 0.0001 * u_NO3,
    y_end = y_start + 0.0001 * u_PO4,
    x_vec = x_start - 0.0003 * u_NO3,
    y_vec = y_start - 0.0003 * u_PO4
  )



ggplot() +
  
  geom_segment(aes(
    x = R_summary %>% filter(ploidy=="dip", nutrient=="NO3") %>% pull(median), xend = 3e-3,
    y = R_summary %>% filter(ploidy=="dip", nutrient=="PO4") %>% pull(median), yend = R_summary %>% filter(ploidy=="dip", nutrient=="PO4") %>% pull(median)),
    color = "dodgerblue", linewidth = 1) +
  
  geom_segment(aes(
    x = R_summary %>% filter(ploidy=="dip", nutrient=="NO3") %>% pull(median), xend = R_summary %>% filter(ploidy=="dip", nutrient=="NO3") %>% pull(median),
    y = R_summary %>% filter(ploidy=="dip", nutrient=="PO4") %>% pull(median), yend = 1.5e-4),
    color = "dodgerblue", linewidth = 1) +
  
  geom_segment(aes(
    x = R_summary %>% filter(ploidy=="tet", nutrient=="NO3") %>% pull(median), xend = 3e-3,
    y = R_summary %>% filter(ploidy=="tet", nutrient=="PO4") %>% pull(median), yend = R_summary %>% filter(ploidy=="tet", nutrient=="PO4") %>% pull(median)),
    color = "deeppink", linewidth = 1) +
  
  geom_segment(aes(
    x = R_summary %>% filter(ploidy=="tet", nutrient=="NO3") %>% pull(median), xend = R_summary %>% filter(ploidy=="tet", nutrient=="NO3") %>% pull(median),
    y = R_summary %>% filter(ploidy=="tet", nutrient=="PO4") %>% pull(median), yend = 1.5e-4),
    color = "deeppink", linewidth = 1) +
  
  geom_segment(aes(x = R_summary %>% filter(ploidy=="dip", nutrient=="NO3") %>% pull(p10),xend = dip_N_p90  <- R_summary %>% filter(ploidy=="dip", nutrient=="NO3") %>% pull(p90),y = R_summary %>% filter(ploidy=="dip", nutrient=="PO4") %>% pull(median),yend = R_summary %>% filter(ploidy=="dip", nutrient=="PO4") %>% pull(median)),color = "darkblue", linewidth = 2, alpha = 0.5) +
  geom_segment(aes(x = R_summary %>% filter(ploidy=="dip", nutrient=="NO3") %>% pull(median),xend = R_summary %>% filter(ploidy=="dip", nutrient=="NO3") %>% pull(median),y = R_summary %>% filter(ploidy=="dip", nutrient=="PO4") %>% pull(p10),yend = R_summary %>% filter(ploidy=="dip", nutrient=="PO4") %>% pull(p90)),color = "darkblue", linewidth = 2, alpha = 0.5) +
  
  geom_segment(aes(x = R_summary %>% filter(ploidy=="tet", nutrient=="NO3") %>% pull(p10), xend = R_summary %>% filter(ploidy=="tet", nutrient=="NO3") %>% pull(p90),y = R_summary %>% filter(ploidy=="tet", nutrient=="PO4") %>% pull(median),yend = R_summary %>% filter(ploidy=="tet", nutrient=="PO4") %>% pull(median)),color = "deeppink4", linewidth = 2, alpha = 0.5) +
  geom_segment(aes(x = R_summary %>% filter(ploidy=="tet", nutrient=="NO3") %>% pull(median),xend = R_summary %>% filter(ploidy=="tet", nutrient=="NO3") %>% pull(median),y = R_summary %>% filter(ploidy=="tet", nutrient=="PO4") %>% pull(p10),yend = R_summary %>% filter(ploidy=="tet", nutrient=="PO4") %>% pull(p90)),color = "deeppink4", linewidth = 2, alpha = 0.5) +
  geom_segment(
    data = arrow_df,
    aes(x = x_start, y = y_start,
        xend = x_end, yend = y_end,
        color = ploidy),
    linewidth = 1,
    arrow = arrow(length = unit(0.25, "cm"))
  ) +
  scale_colour_manual(values = c("dip" = "dodgerblue", "tet" = "deeppink")) +
  geom_segment(
    data = arrow_df,
    aes(x = x_start, y = y_start,
        xend = x_vec, yend = y_vec,
        color = ploidy),
    linewidth = 0.5, linetype = 2) +
  #geom_density2d(data = R_points %>% filter(ploidy == "dip"),aes(x = R_min_NO3, y = R_min_PO4),colour = "royalblue3",alpha = 0.2) +
  #geom_density2d(data = R_points %>% filter(ploidy == "tet"),aes(x = R_min_NO3, y = R_min_PO4),colour = "deeppink3",alpha = 0.2) +

  geom_point(data = R_points %>% filter(ploidy == "dip"),aes(x = R_min_NO3, y = R_min_PO4),color = "dodgerblue", alpha = 0.1, size = 0.5) +
  
  geom_point(data = R_points %>% filter(ploidy == "tet"),aes(x = R_min_NO3, y = R_min_PO4),color = "deeppink", alpha = 0.1, size = 0.5) +
  xlab("NO3 (mg/L)") +
  ylab("PO4 (mg/L)") +
  theme_bw()

ggsave("Rstar_consumption_vectors100.png", width = 6, height = 5)

```

```{r}
summary(posterior)
parameters_simple <- summary(posterior) %>%
  dplyr::select(variable, mean)

summary(posteriorVK)
parameters_simpleVK <- summary(posteriorVK) %>%
  dplyr::select(variable, mean)

VK_ <- parameters_simpleVK %>%
  filter(grepl("logV_", variable) | grepl("logK_", variable)) %>%
  mutate(
    ploidy = ifelse(grepl("ploidydip", variable), "dip", "tet"),
    medium = ifelse(grepl("nutrientN", variable), "N", "P"),
    param  = ifelse(grepl("logV_", variable), "logV", "logK")
  ) %>%
  select(ploidy, medium, param, mean) %>%
  tidyr::pivot_wider(names_from = param, values_from = mean) %>%
  mutate(
    V = exp(logV),
    K   = exp(logK)
  )

R_ <- parameters_simple %>%
  filter(grepl("b_r_", variable) | grepl("b_qmin_", variable) | grepl("b_mortality_", variable)) %>%
  mutate(
    ploidy = ifelse(grepl("ploidydip", variable), "dip", "tet"),
    medium = ifelse(grepl("mediumN", variable), "N", "P"),
    param = case_when(
      grepl("b_r_", variable) ~ "r",
      grepl("b_qmin_", variable) ~ "qmin",
      grepl("b_mortality_", variable) ~ "mortality")
  ) %>%
  select(ploidy, medium, param, mean) %>%
  tidyr::pivot_wider(names_from = param, values_from = mean)

R_etoile <- R_ %>%
  left_join(VK_, by = c("ploidy", "medium")) %>%
  left_join(
    nu %>% select(ploidy, nutrient, condition),
    by = c("ploidy" = "ploidy", "medium" = "nutrient")) %>%
  mutate(
    R_min = (r * qmin * K * mortality) /
            (V * (r - mortality) - r * qmin * mortality)) %>%
  mutate(nutrient = medium) %>%
  select(-condition)

R_etoile_dip <- R_etoile %>%
  filter(ploidy == "dip") %>%
  select(ploidy, nutrient, R_min) %>%
  distinct()
R_etoile_tet <- R_etoile %>%
  filter(ploidy == "tet") %>%
  select(ploidy, nutrient, R_min) %>%
  distinct()


nutrient.ratio <- bind_rows(
  R_etoile_dip %>%
    pivot_wider(names_from = medium, values_from = R_min) %>%
    mutate(ratio = N / P),
  R_etoile_tet %>%
    pivot_wider(names_from = medium, values_from = R_min) %>%
    mutate(ratio = N / P)
)

ggplot(nutrient.ratio) +
  geom_vline(aes(xintercept = ratio, color = ploidy),
    linewidth = 0.7, linetype = "dashed") +
  geom_jitter(data = nu, aes(x = ratio, y = ploidy, color = ploidy),
            height = 0.2, width = 0) +
  scale_color_manual(values = c("dip" = "royalblue", "tet" = "deeppink3")) +
  annotate("text", x = 35, y = 1, xend = 36, label = paste0("Phosphate limited"), size = 3.5) +
  annotate("text", x = 28.8, y = 1, xend = 28, label = paste0("Nitrate limited"), size = 3.5) +
  annotate("text", x = 31.9, y = 1, label = paste0("Coexistence zone"), size = 3.5) +
  scale_x_log10(breaks = c(0.1, 10, 100)) +
  labs(x = "NO3/PO4") + 
  theme_minimal()


#how big the vectors, coexistence dashed lines should be
sc <- 0.05

nu0 <- nu %>%
  select(ploidy, medium, nutrient, S0) %>%
  filter(ploidy == c("dip", "tet"))
  
consumption <- R_etoile %>%
  left_join(nu0) %>%
  mutate(
    condition = interaction(ploidy, medium),
    N0 = 1,
    S0 = mean(S0, na.rm = T),
    S_t = (N0 * V * (S0 / (K + S0)))) %>%
  select(ploidy, nutrient, medium, condition, S_t) %>%
  group_by(condition) %>%
  mutate(meanS_t = mean(S_t, na.rm = TRUE)) %>%
  ungroup()

consumption_vectors <- consumption %>%
  mutate(
    S_N = ifelse(nutrient == "N", meanS_t, meanS_t[nutrient == "N" & ploidy == "dip"]),
    S_P = ifelse(nutrient == "P", meanS_t, meanS_t[nutrient == "P" & ploidy == "tet"]),
    norm = round(sqrt(S_N^2 + S_P^2), 5)
    ) %>%
  select(-S_t, -meanS_t)%>%
  mutate(norm = ifelse(ave(norm, norm, FUN = length) >= 4, NA, norm)) %>%
  mutate(
    u_N = S_N / norm,
    u_P = S_P / norm,
    
#if the vectors are wrong, inverse tet with dip
    x_ori = R_etoile_tet %>% filter(nutrient == "N") %>% pull(R_min),
    y_ori = R_etoile_dip %>% filter(nutrient == "P") %>% pull(R_min),
    x_end_vector = x_ori - 0.2 * sc * u_N,
    y_end_vector = y_ori - 0.2 * sc * u_P,
    x_coex = x_ori + sc * u_N,
    y_coex = y_ori + sc * u_P)

theta <- consumption_vectors %>%
  filter(ploidy %in% c("dip", "tet")) %>%
  select(condition, ploidy, u_N, u_P) %>%
  group_by(condition, ploidy) %>% 
  summarise(
    u_N = mean(u_N, na.rm = TRUE),
    u_P = mean(u_P, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  tidyr::pivot_wider(
    names_from = ploidy,
    values_from = c(u_N, u_P)
  ) %>%
  filter(
    !is.na(u_N_dip), !is.na(u_N_tet),
    !is.na(u_P_dip), !is.na(u_P_tet)
  ) %>%
  mutate(
    theta_rad = acos(u_N_dip * u_N_tet + u_P_dip * u_P_tet),
    theta_deg = theta_rad * 180/pi
  )




shade_df <- consumption_vectors %>%
  mutate(
    x_end = x_ori + sc * u_N,
    y_end = y_ori + sc * u_P
  )

poly_df <- shade_df %>%
  filter(ploidy %in% c("dip", "tet")) %>%
  arrange(ploidy) %>%
  reframe(
    x = c(x_ori[38], x_end[38], x_end[39]),
    y = c(y_ori[38], y_end[38], y_end[39])
  )

p_ <- ggplot() +
  #geom_polygon(data = poly_df, aes(x, y),
   # fill = "grey80", alpha = 0.5) +
  
  geom_segment(aes(
    x = R_etoile_dip %>% filter(nutrient == "N") %>% pull(R_min),
    xend = 3,
    y = R_etoile_dip %>% filter(nutrient == "P") %>% pull(R_min),
    yend = R_etoile_dip %>% filter(nutrient == "P") %>% pull(R_min)
  ), color = "dodgerblue", linewidth = 1) +

geom_segment(aes(
    x = R_etoile_dip %>% filter(nutrient == "N") %>% pull(R_min),
    xend = R_etoile_dip %>% filter(nutrient == "N") %>% pull(R_min),
    y = R_etoile_dip %>% filter(nutrient == "P") %>% pull(R_min),
    yend = 0.03
  ), color = "dodgerblue", linewidth = 1) +

geom_segment(aes(
    x = R_etoile_tet %>% filter(nutrient == "N") %>% pull(R_min),
    xend = 3,
    y = R_etoile_tet %>% filter(nutrient == "P") %>% pull(R_min),
    yend = R_etoile_tet %>% filter(nutrient == "P") %>% pull(R_min)
  ), color = "deeppink", linewidth = 1) +

geom_segment(aes(
    x = R_etoile_tet %>% filter(nutrient == "N") %>% pull(R_min),
    xend = R_etoile_tet %>% filter(nutrient == "N") %>% pull(R_min),
    y = R_etoile_tet %>% filter(nutrient == "P") %>% pull(R_min),
    yend = 0.03
  ), color = "deeppink", linewidth = 1) +

  #geom_segment(
   # data = shade_df,
    #aes(x = x_ori, y = y_ori,
     #   xend = x_end_vector, yend = y_end_vector,
      #  color = ploidy),
    #linewidth = 0.8, arrow = arrow(length = unit(0.25, "cm"))) +
  
  #geom_segment(
   # data = shade_df,
    #aes(x = x_ori, y = y_ori, xend = x_coex, yend = y_coex,
     #   color = ploidy),
    #linewidth = 0.8, linetype = "dashed") +
  
  scale_color_manual(values = c("dip" = "royalblue", "tet" = "deeppink3")) +
  
  #annotate("text", 
   #        label = paste0("θ = ", round(theta["theta_deg"], 2), "°   cos(θ) = ", #round(cos(theta["theta_rad"]), 3)), size = 3.5, x = 1, y = 0.02) +
  
  labs(x = "NO3", y = "PO4") +
  theme_bw()

p_

```

# Test

```{r }
# Time points
times <- seq(from = 0, to = 50, length.out = 100)

# Function T(t)
T_fun <- function(t, N0, V, r, S0, K) {
  (N0 * V / r) * (exp(r * t) - 1) + S0 + K * log(S0)
}

# Solve S + K ln S = T(t)
solve_S <- function(Tt, K) {
  f <- function(S) S + K * log(S) - Tt
  
  lower <- .Machine$double.eps
  upper <- max(1, Tt) * 1000
  
  uniroot(f, interval = c(lower, upper))$root
}


# Compute S(t) and consumption C(t)
compute_consumption <- function(N0, V, r, K, S0, times) {
  T_vals <- T_fun(times, N0, V, r, S0, K)
  S_vals <- sapply(T_vals, solve_S, K = K)
  C_vals <- N0 * exp(r * times) * V * (S_vals / (K + S_vals))
  
  tibble(
    t   = times,
    S_t = S_vals,
    C_t = C_vals
  )
}


# Apply to your ploidy × medium parameters

R_etoile <- R_ %>%
  left_join(VK_, by = c("ploidy", "medium")) %>%
  left_join(
    conditions %>% select(ploidy, nutrient),
    by = c("ploidy" = "ploidy", "medium" = "nutrient")) %>%
  mutate(
    R_min = (r * qmin * K) /
            (V * (r) - r * qmin)
  )

consumption_vectors <- R_etoile %>%
  mutate(
    S0 = ifelse(medium == "N", 550, 50),
    out = pmap(
      list(N0, V, r, K, S0),
      ~ compute_consumption(..1, ..2, ..3, ..4, ..5, times)
    )
  ) %>%
  unnest(out) %>%
  mutate(
    S_N = ifelse(medium == "N", C_t, NA_real_),
    S_P = ifelse(medium == "P", C_t, NA_real_)
  ) %>%
  select(ploidy, medium, t, S_N, S_P) %>%
  pivot_wider(
    names_from = medium,
    values_from = c(S_N, S_P)
  ) %>%
  rename(
    S_N = S_N_N,
    S_P = S_P_P
  ) %>%
  mutate(
    norm = sqrt(S_N^2 + S_P^2),
    u_N = S_N / norm,
    u_P = S_P / norm,
    x_start = R_dip["N"],
    y_start = R_tet["P"],
    x_end   = x_start - 0.1 * u_N,
    y_end   = y_start - 0.1 * u_P,
    x_dash  = x_start + 0.1 * u_N,
    y_dash  = y_start + 0.1 * u_P
    ) %>%
  group_by(t) %>%
  mutate(
    # dip vector
    v1_N = u_N[ploidy == "dip"],
    v1_P = u_P[ploidy == "dip"],

    # tet vector
    v2_N = u_N[ploidy == "tet"],
    v2_P = u_P[ploidy == "tet"],

    # angle
    theta = acos(v1_N * v2_N + v1_P * v2_P),
    theta_deg = theta * 180 / pi
  ) %>%
  ungroup()


library(gganimate)

p_anim <- ggplot(consumption_vectors) +
  
  # Dip ZNGI
  geom_segment(
    aes(x = R_dip["N"], xend = 0.25,
        y = R_dip["P"], yend = R_dip["P"]),
    inherit.aes = FALSE,
    color = "dodgerblue", linewidth = 1
  ) +
  geom_segment(
    aes(x = R_dip["N"], xend = R_dip["N"],
        y = R_dip["P"], yend = 0.1),
    inherit.aes = FALSE,
    color = "dodgerblue", linewidth = 1
  ) +
  
  # Tet ZNGI
  geom_segment(
    aes(x = R_tet["N"], xend = 0.25,
        y = R_tet["P"], yend = R_tet["P"]),
    inherit.aes = FALSE,
    color = "deeppink", linewidth = 1
  ) +
  geom_segment(
    aes(x = R_tet["N"], xend = R_tet["N"],
        y = R_tet["P"], yend = 0.1),
    inherit.aes = FALSE,
    color = "deeppink", linewidth = 1
  ) +
  
  # Consumption vectors
  geom_segment(
    aes(x = x_start, y = y_start,
        xend = x_end, yend = y_end,
        color = ploidy),
    linewidth = 1,
    arrow = arrow(length = unit(0.25, "cm"))
  ) +
  
  geom_segment(
    aes(x = x_start, y = y_start,
        xend = x_dash, yend = y_dash,
        color = ploidy),
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  
  scale_color_manual(values = c("dip" = "royalblue", "tet" = "deeppink3")) +
  
  labs(
    x = "NO3 (mmol.L-1)",
    y = "PO4 (mmol.L-1)",
    title = "Consumption vectors at time t = {closest_state}"
  ) +
  
  theme_bw() +
  
  transition_states(t, transition_length = 2, state_length = 1)

anim <- gganimate::animate(
  p_anim,
  nframes = 120,
  fps = 20,
  width = 800,
  height = 600
)

gganimate::anim_save(
  "C:/Users/tozah/Desktop/Internship Ghent University/Protocols & Experiments/consumption_vectors.gif",
  animation = anim
)

# data.frame to be filled
wf_df <- data.frame()

# effective population sizes
sizes <- c(200, 1000, 5000)

# starting allele frequencies
starting_p <- c(.1, .5)

# number of generations
n_gen <- 1000

# number of replicates per simulation
n_reps <- 50

# run the simulations
for(N in sizes){
  for(p in starting_p){
    p0 <- p
    for(j in 1:n_gen){
      X <- rbinom(n_reps, 2*N, p)
      p <- X / (2*N)
      rows <- data.frame(replicate = 1:n_reps, N = rep(N, n_reps), 
                         gen = rep(j, n_reps), p0 = rep(p0, n_reps), 
                         p = p)
      wf_df <- bind_rows(wf_df, rows)
    }
  }
}

# plot it up!
p <- ggplot(wf_df, aes(x = gen, y = p, group = replicate)) +
     geom_path(alpha = .5) + facet_grid(N ~ p0) + guides(colour=FALSE)
p
```


# Finding equilibriums in the Droop model

```{r }
# state vector (example equilibrium)
state <- c(
  Ndip = 1,
  Ntet = 1,
  QdN  = 0.5,
  QtN  = 0.5,
  QdP  = 0.5,
  QtP  = 0.5,
  SN   = 550,
  SP   = 50
)

# parameter list
p <- readRDS("C:/Users/tozah/Desktop/Internship Ghent University/Protocols & Experiments/p.RDS")

jacobian <- function(x,p){
  Nd<-x[1]; Nt<-x[2]; QdN<-x[3]; QtN<-x[4]; QdP<-x[5]; QtP<-x[6]; SN<-x[7]; SP<-x[8]
  g  <- function(S,K) S/(S+K); gp <- function(S,K) K/(S+K)^2
  J <- matrix(0,8,8)
  J[1,1] <- p$r_dip*(1 - p$q_min_dip_N/QdN) - p$m_dip_N
  J[1,3] <- Nd*p$r_dip*p$q_min_dip_N/QdN^2
  J[2,2] <- p$r_tet*(1 - p$q_min_tet_N/QtN) - p$m_tet_N
  J[2,4] <- Nt*p$r_tet*p$q_min_tet_N/QtN^2
  J[3,3] <- -p$r_dip
  J[3,7] <- p$V_dip_N*gp(SN,p$K_dip_N)
  J[4,4] <- -p$r_tet
  J[4,7] <- p$V_tet_N*gp(SN,p$K_tet_N)
  J[5,5] <- -p$r_dip
  J[5,8] <- p$V_dip_P*gp(SP,p$K_dip_P)
  J[6,6] <- -p$r_tet
  J[6,8] <- p$V_tet_P*gp(SP,p$K_tet_P)
  J[7,1] <- -p$V_dip_N*g(SN,p$K_dip_N)
  J[7,2] <- -p$V_tet_N*g(SN,p$K_tet_N)
  J[7,7] <- -(Nd*p$V_dip_N*gp(SN,p$K_dip_N) + Nt*p$V_tet_N*gp(SN,p$K_tet_N))
  J[8,1] <- -p$V_dip_P*g(SP,p$K_dip_P)
  J[8,2] <- -p$V_tet_P*g(SP,p$K_tet_P)
  J[8,8] <- -(Nd*p$V_dip_P*gp(SP,p$K_dip_P) + Nt*p$V_tet_P*gp(SP,p$K_tet_P))
  J
}

J  <- jacobian(state, p)
tr <- sum(diag(J))
dt <- det(J)

tr
dt

tr_range <- seq(-2, 2, length.out = 500)
det_curve <- tr_range^2 / 4

df_curve <- data.frame(
  det = det_curve,
  tr  = tr_range
)

ggplot() +
  geom_line(data = df_curve, aes(tr, det), color = "dodgerblue", linewidth = 1) +
  geom_point(aes(tr, dt), size = 3, color = "blue") +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.6) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.6) +
  labs(x = "trace(J)", y = "det(J)") +
  theme_minimal()



```

```{python}

x = 0
y = 2

if(x == 0){
  y = 1
} else {
  y = 0
}

print(y)

```